11. Better Simulations and HMM estimations
-------------------------------------
```{r eval=TRUE, echo=FALSE}

source("SharedFunctions.R")

```

#### Simulating and estimating for 10,000 loci
recombination rate = 0.0002 
```{r}
#set.seed(1234455)
start.p<-c(0.25,0.25,0.25,0.25)
recombination.rate<-0.0004
prob.missing=0.1
trans.p<-GetTransitionProb(recombination.rate)
child1.chromatid.mom<-SimulateChromatid(10000,recombination.rate) 
child1.chromatid.dad<-SimulateChromatid(10000,recombination.rate)
old_obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
obs<-GetDiploidGenotypeWithErrors(child1.chromatid.mom,child1.chromatid.dad,0.1)
old_obs[1:10]
obs[1:10]
v.2<-ViterbiWithMissingDataAndError (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad,prob.missing)
 
vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
#vitrowmax.2
simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
estimated<-rownames(trans.p)[vitrowmax.2]
sim.results<-sum(simulated==estimated)

```
The number for true positives from our estimation would be given by. 
```{r}
sum(sim.results)
CountRecombEvents(estimated)
CountRecombEvents(simulated)
simulated[1:10]
estimated[1:10]

```
and the percentage of true positives is `r  sum(sim.results)/10000*100`%

### Simulating with errors
```{r}
ViterbiSimulatorWithErrors <- function(num.markers,num.simulations,start.p,recombination.rate,prob.missing){ 
  sim.results<-c()
  sim.recomb.events<-c()
  est.recomb.events<-c()
  trans.p<-GetTransitionProb(recombination.rate)
  for (i in 1:num.simulations){
    child1.chromatid.mom<-SimulateChromatid(num.markers,recombination.rate) 
    child1.chromatid.dad<-SimulateChromatid(num.markers,recombination.rate)
    #obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
    obs<-GetDiploidGenotypeWithErrors(child1.chromatid.mom,child1.chromatid.dad,prob.missing)
    v.2<-ViterbiWithMissingDataAndError(obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad,prob.missing)
    vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
    simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
    estimated<-rownames(trans.p)[vitrowmax.2]
    sim.results[i]<-sum(simulated==estimated)
    sim.recomb.events[i]<-CountRecombEvents(simulated);
    est.recomb.events[i]<-CountRecombEvents(estimated);
  }
  return(list(sim.results=sim.results,sim.recomb.events=sim.recomb.events,est.recomb.events=est.recomb.events))
}  
 
```
And we can run simulations
We will peform a simulation and estimation and count simulated and estimated recombination events.
```{r}
simul.results<-ViterbiSimulatorWithErrors(10000,200,start.p,recombination.rate,prob.missing)

summary(simul.results)

summary(reg1<-lm(simul.results$est.recomb.events~simul.results$sim.recomb.events))
plot(simul.results$est.recomb.events~simul.results$sim.recomb.events,ylab="estimated",xlab="simulated",main="Number of recombination events\n10,000 loci, r=0.0002")
abline(reg1)


```

We can see that although we are overestimating recombination events, they are proportional to the simulted number of recombination events.
Now let's test the same but with 10,000 markers and recombination rate = 0.0002 (200 simulations).


#### Author
Luis Avila (lmavila@gmail.com)