4. The Viterbi Algorithm
-----------------------------

#### First implementation
This is just a test of a viterbi implementation
with a fix Emission matrix for 5 observations.
Adapted from (Krijnen 2009).
```{r eval=TRUE, echo=FALSE}

source("SharedFunctions.R")


```


```{r}

start.p<-c(0.4,0.2,0.2,0.2) #The starting probabilities chosen arbitrarily 
emit.p<-GetEmissionProb(0,0,1,0) # For this example we are having just one emission matrix, not 16 
                                 # to keep it simple for now we are not using the parent genotypes or phasing.
#obs<-child1.genotype
obs<-c(0,0,1,1,2)
trans.p<-GetTransitionProb(0.2) # Getting the transition probabilities matrix

Viterbi <- function(obs,start.p, trans.p, emit.p) {
  #initializing
  v <- matrix(NA, nr=length(obs), nc=dim(trans.p)[1]) ## we create a 10x4 empty matrix (10 obs, 4 hidden states)
  for(i in 1:4){ #for each of the hidden states we calculate the initial probabilities
    v[1,i]=start.p[i]*emit.p[i,obs[1]+1]
  }
  
  for(i in 2:length(obs)) {# from observation 2 to t
    for (l in 1:dim(trans.p)[1]) { #from state 1 to N (4 states)
      v[i,l] <- emit.p[l,obs[i]+1] * max(v[(i-1),] * trans.p[l,])
    }
  }   
      
return(v) 
}


v<-Viterbi (obs,start.p, trans.p, emit.p)
v
```
With the matrix v calculated, the state path given us the maximum 
probability can be found by the maximum values per row, with this function.

```{r}
vitrowmax <- apply(v, 1, function(x) which.max(x))
vitrowmax
```
#### Viterbi: Second implementation
Now we are going to adjust our implementation of the Viterbi algorithm 
to use the information we have about the parents phasing.

```{r}
child1.chromatid.mom<-SimulateChromatid()
child1.chromatid.mom
child1.chromatid.dad<-SimulateChromatid()
child1.chromatid.dad

### for our observation in the child
obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
obs


start.p<-c(0.25,0.25,0.25,0.25);

trans.p<-GetTransitionProb(0.3)

Viterbi2 <- function(obs,start.p, trans.p,mom.chromatid,dad.chromatid) {
  #initializing
  v <- matrix(NA, nr=length(obs), nc=dim(trans.p)[1])
  emit.p<-GetEmissionProb(mom.chromatid$maternal[1],
                          mom.chromatid$paternal[1],
                          dad.chromatid$maternal[1],
                          dad.chromatid$paternal[1])
  for(i in 1:4){
    v[1,i]=start.p[i]*emit.p[i,obs[1]+1]
  }
  
  for(i in 2:length(obs)) {# from observation 2 to t
    emit.p<-GetEmissionProb(mom.chromatid$maternal[i],
                            mom.chromatid$paternal[i],
                            dad.chromatid$maternal[i],
                            dad.chromatid$paternal[i])
    for (l in 1:dim(trans.p)[1]) { #from state 1 to N (4 states)
      v[i,l] <- emit.p[l,obs[i]+1] * max(v[(i-1),] * trans.p[l,])
    }
  }   
  
  return(v) 
}



v.2<-Viterbi2 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
v.2
```
And to get the parent path

```{r}

vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
vitrowmax.2
```

And in order to compare our simulated states with the estimated states obtained using our
HHMmodel, the observed genotype of the simulated progeny and the genotype and phasing of
the simulated parents.
 
```{r}
simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
estimated<-rownames(trans.p)[vitrowmax.2]
simulated
estimated
simulated==estimated

```

The model seems to be working with the simulated data, some predictions are right
and some are wrong. To better assess the accuracy of the model we will run simulations
with more markers and different recombination rates.

#### References
Krijnen WP. 2009. Applied Statistics for Bioinformatics using R. GNU Free Document License.

#### Author
Luis Avila (lmavila@gmail.com)