5. Simulations and HMM estimations
----------------------------------
```{r eval=TRUE, echo=FALSE}

source("SharedFunctions.R")

```

#### Simulating and estimating for 100 loci
recombination rate = 0.3 
```{r}
start.p<-c(0.25,0.25,0.25,0.25)
recombination.rate<-0.3
trans.p<-GetTransitionProb(recombination.rate)
child1.chromatid.mom<-SimulateChromatid(100,recombination.rate) #100 markers
child1.chromatid.dad<-SimulateChromatid(100,recombination.rate)
obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
v.2<-Viterbi2 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
vitrowmax.2
simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
estimated<-rownames(trans.p)[vitrowmax.2]
sim.results<-sum(simulated==estimated)
```
The number for true positives from our estimation would be given by. 
```{r}
sum(sim.results)
```
and the percentage of true positives is `r sum(sim.results)/100*100`%

#### Simulating for 10,000 loci and recombination rate r=0.3
```{r}
child1.chromatid.mom<-SimulateChromatid(10000,recombination.rate) #100 markers
child1.chromatid.dad<-SimulateChromatid(10000,recombination.rate)
obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
v.2<-Viterbi2 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
estimated<-rownames(trans.p)[vitrowmax.2]
sim.results<-sum(simulated==estimated)
sum(sim.results)
```
and the percentage of true positives is `r (sum(sim.results)/10000)*100`%.
Much lower than when we simulated for 100 loci.
Let's look at some values of our matrix returned by Viterbi2 to see what is the problem.
```{r}
v.2[1:5,]
v.2[1500:1505,]
v.2[5000:5005,]
v.2[9000:9005,]
```
As you can see, the values get smaller to the point R does not have a representation and shows 0.
To prevent this we can modify our Viterbi2 algorythm to use the log2 of the probabilities instead.
```{r}
Viterbi3 <- function(obs,start.p, trans.p,mom.chromatid,dad.chromatid) {
  #initializing
  v <- matrix(NA, nr=length(obs), nc=dim(trans.p)[1])
  emit.p<-GetEmissionProb(mom.chromatid$maternal[1],
                          mom.chromatid$paternal[1],
                          dad.chromatid$maternal[1],
                          dad.chromatid$paternal[1])
  emit.p<-emit.p+10^(-12)                        
  for(i in 1:4){
    v[1,i]=log(start.p[i],2)+log(emit.p[i,obs[1]+1],2)
  }
  
  for(i in 2:length(obs)) {# from observation 2 to t
     emit.p<-GetEmissionProb(mom.chromatid$maternal[i],
                             mom.chromatid$paternal[i],
                             dad.chromatid$maternal[i],
                             dad.chromatid$paternal[i])
     for (l in 1:dim(trans.p)[1]) { #from state 1 to N (4 states)
       v[i,l] <- log(emit.p[l,obs[i]+1]+10^(-12),2) + max(v[(i-1),] + log( trans.p[l,],2))
     }
  }   
  
  return(v) 
}


```
You will notice that I am adding +10^(-12) before using the value from the emission matrix.
We are doing that because we don't want to calculate the log2 of 0 which would be -Inf.

Using the same simulated dat we can run our new Viterbi3 funcion.

```{r}
child1.chromatid.mom<-SimulateChromatid(10000,recombination.rate) #100 markers
child1.chromatid.dad<-SimulateChromatid(10000,recombination.rate)
obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
v.2<-Viterbi3 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
estimated<-rownames(trans.p)[vitrowmax.2]
sim.results<-sum(simulated==estimated)
sum(sim.results)
```
and the percentage of true positives is `r (sum(sim.results)/10000)*100`%.
Closer to what we got when simulated for 100 loci.
Now that our Viterbi function is more robust we can do more simulations.

#### A lot of simulations

```{r}
start.p<-c(0.25,0.25,0.25,0.25)

ViterbiSimulator <- function(num.markers,num.simulations,start.p,recombination.rate){ 
  sim.results<-c()
  trans.p<-GetTransitionProb(recombination.rate)
  for (i in 1:num.simulations){
    child1.chromatid.mom<-SimulateChromatid(num.markers,recombination.rate) #100 markers
    child1.chromatid.dad<-SimulateChromatid(num.markers,recombination.rate)
    obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
    v.2<-Viterbi3 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
    vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
    simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
    estimated<-rownames(trans.p)[vitrowmax.2]
    sim.results[i]<-sum(simulated==estimated)
  }
  return(sim.results)
}  

result.100.03<-ViterbiSimulator(100,100,start.p,0.3);
hist(result.100.03,main="True positives in 100 simulations with 100 markers\n r=0.3",xlab="true positives")
result.100.02<-ViterbiSimulator(100,100,start.p,0.2);
hist(result.100.02,main="True positives in 100 simulations with 100 markers\n r=0.2",xlab="true positives")
result.100.01<-ViterbiSimulator(100,100,start.p,0.1);
hist(result.100.01,main="True positives in 100 simulations with 100 markers\n r=0.1",xlab="true positives")
result.500.03<-ViterbiSimulator(500,100,start.p,0.3)
result.500.02<-ViterbiSimulator(500,100,start.p,0.2)
result.500.01<-ViterbiSimulator(500,100,start.p,0.1)

hist(result.500.03,main="True positives in 100 simulations with 500 markers\n r=0.3",xlab="true positives")
hist(result.500.01,main="True positives in 100 simulations with 500 markers\n r=0.1",xlab="true positives")

```
```{r eval=TRUE, echo=FALSE}
result.300.03<-ViterbiSimulator(300,100,start.p,0.3)
result.300.02<-ViterbiSimulator(300,100,start.p,0.2)
result.300.01<-ViterbiSimulator(300,100,start.p,0.1)
result.10000.03<-ViterbiSimulator(10000,100,start.p,0.3)
result.10000.02<-ViterbiSimulator(10000,100,start.p,0.2)
result.10000.01<-ViterbiSimulator(10000,100,start.p,0.1)




```

#### Percent mean true positives for estimations of different recombination rates and number of loci

x               |100 loci      |300 loci  |500 loci      |10,000 loci|
---------------:|:--------:|:--------:|:--------:|:----------:|
**r=0.3**       | `r mean(result.100.03)/100*100`|`r median(result.300.03)/300*100`|`r min(result.500.03)/500*100`| `r mean(result.10000.03)/10000*100`| 
**r=0.2**       | `r mean(result.100.02)/100*100`|`r median(result.300.02)/300*100`|`r min(result.500.02)/500*100`|`r mean(result.10000.02)/10000*100`|      
**r=0.1**       |`r mean(result.100.01)/100*100`|`r median(result.300.01)/300*100`|`r min(result.500.01)/500*100`|  `r mean(result.10000.01)/10000*100`|
     

From these results it is clear that our estimation accuracy increases with smaller recombination rates and it appears to be independent from the number of loci.

#### Author
Luis Avila (lmavila@gmail.com)