7. Counting recombination events
----------------------------------
```{r eval=TRUE, echo=FALSE}

source("SharedFunctions.R")

```
We can count the recombination events from the state path data with this function.
```{r}
CountRecombEvents<-function(parent.path){
  current<-parent.path[1];
  counter<-0;
  for(val in parent.path){
    if(val!=current){
      counter<-counter+1
      current<-val
    }
  }
  return (counter)
}

```
We will peform a simulation and estimation and count simulated and estimated recombination events.
```{r}
  start.p<-c(0.25,0.25,0.25,0.25)
  recombination.rate<-0.001
  trans.p<-GetTransitionProb(recombination.rate)
  child1.chromatid.mom<-SimulateChromatid(10000,recombination.rate) 
  child1.chromatid.dad<-SimulateChromatid(10000,recombination.rate)
  obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
  v.2<-Viterbi2 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
  vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
  simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
  estimated<-rownames(trans.p)[vitrowmax.2]
  sim.results<-sum(simulated==estimated)
```
The number for true positives from our estimation would be given by. 
```{r}
  sum(sim.results)
```
and the percentage of true positives is `r sum(sim.results)/10000*100`%

And counting recombination events
```{r}
   CountRecombEvents(simulated)
   CountRecombEvents(estimated)
   CountRecombEvents(estimated)-CountRecombEvents(simulated)
   10000-sim.results
```
Another simulation using the same recombination rate
```{r}
  child1.chromatid.mom<-SimulateChromatid(10000,recombination.rate) 
  child1.chromatid.dad<-SimulateChromatid(10000,recombination.rate)
  obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
  v.2<-Viterbi2 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
  vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
  simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
  estimated<-rownames(trans.p)[vitrowmax.2]
  sim.results<-sum(simulated==estimated)
  sim.results
  CountRecombEvents(simulated)
  CountRecombEvents(estimated)
  CountRecombEvents(estimated)-CountRecombEvents(simulated)
  10000-sim.results
```
I appears wrong estimations from our model make us overestimate the number of recombination events.
Let's update our simulator to report the numer of recombination events per simulation in 
addition to the number of correct estimation of parental contribuitions.
```{r}
ViterbiSimulator2 <- function(num.markers,num.simulations,start.p,recombination.rate){ 
  sim.results<-c()
  sim.recomb.events<-c()
  est.recomb.events<-c()
  trans.p<-GetTransitionProb(recombination.rate)
  for (i in 1:num.simulations){
    child1.chromatid.mom<-SimulateChromatid(num.markers,recombination.rate) #100 markers
    child1.chromatid.dad<-SimulateChromatid(num.markers,recombination.rate)
    obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
    v.2<-Viterbi3 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
    vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
    simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
    estimated<-rownames(trans.p)[vitrowmax.2]
    sim.results[i]<-sum(simulated==estimated)
    sim.recomb.events[i]<-CountRecombEvents(simulated);
    est.recomb.events[i]<-CountRecombEvents(estimated);
  }
  return(list(sim.results=sim.results,sim.recomb.events=sim.recomb.events,est.recomb.events=est.recomb.events))
}  

simul.results<-ViterbiSimulator2(1000,100,c(0.25,0.25,0.25,0.25),0.01)
summary(simul.results)

summary(reg1<-lm(simul.results$est.recomb.events~simul.results$sim.recomb.events))
plot(simul.results$est.recomb.events~simul.results$sim.recomb.events)
abline(reg1)


```

We can see that although we are overestimating recombination events, they are proportional to the simulted number of recombination events.
Now let's test the same but with 10,000 markers and recombination rate = 0.001 (50 simulations).

```{r}
simul.results<-ViterbiSimulator2(10000,50,c(0.25,0.25,0.25,0.25),0.001)
summary(simul.results)
summary(reg1<-lm(simul.results$est.recomb.events~simul.results$sim.recomb.events))
plot(simul.results$est.recomb.events~simul.results$sim.recomb.events)
abline(reg1)
```

#### Author
Luis Avila (lmavila@gmail.com)