8. Estimating recombination rate
----------------------------------
```{r eval=TRUE, echo=FALSE}

source("SharedFunctions.R")

```
#### First we simulate for 10,000 loci and r=0.001

```{r}
  start.p<-c(0.25,0.25,0.25,0.25)
  recombination.rate<-0.001 #simulating with r=0.001

  trans.p<-GetTransitionProb(recombination.rate)
  child1.chromatid.mom<-SimulateChromatid(10000,recombination.rate) 
  child1.chromatid.dad<-SimulateChromatid(10000,recombination.rate)
  simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
  obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
```  
#### And now we try to estimate using an arbitrary r=0.4 in our transition matrix
```{r}
  

  trans.p<-GetTransitionProb(0.4); #we will estimate with r=0.4
  v.2<-Viterbi3 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
  vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
  estimated<-rownames(trans.p)[vitrowmax.2]
  CountRecombEvents(simulated)
  CountRecombEvents(estimated)
  
```
And our estimated recombination rate could be calculated like this:
```{r}
  CountRecombEvents(estimated)/(10000*2)
  
  
```

Now testing different initial recombination rates
```{r}
  rates<-c(0.5,0.4,0.3,0.3,0.2,0.1,0.01,0.001)
  estimated.rate<-c()
  recomb.count<-c()
  index<-1;
  for(rate in rates){
    trans.p<-GetTransitionProb(rate); 
    v.2<-Viterbi3 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
    vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
    estimated<-rownames(trans.p)[vitrowmax.2]
    recomb.count[index]<-CountRecombEvents(estimated)
    estimated.rate[index]<- recomb.count[index]/(10000*2);
    index<-index+1
  }
  
  rates
  estimated.rate
  recomb.count
  

    
  plot(rates,estimated.rate,xlab="arbitrary recomb. rate given to transition matrix",ylab="estimated recomb. rate",main="r.rate estimations with different starting r.rate\n on one simulation with 10,000 loci and r=0.001")
  
```

#### Now we simulate with 10,000 loci and r=0.01

```{r}
  recombination.rate<-0.01 #simulating with r=0.001
  trans.p<-GetTransitionProb(recombination.rate)
  child1.chromatid.mom<-SimulateChromatid(10000,recombination.rate) 
  child1.chromatid.dad<-SimulateChromatid(10000,recombination.rate)
  simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
  obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
  
```

#### And now we try to estimate using an arbitrary r=0.4 in our transition matrix
```{r}
  

  trans.p<-GetTransitionProb(0.4); #we will estimate with r=0.4
  v.2<-Viterbi3 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
  vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
  estimated<-rownames(trans.p)[vitrowmax.2]
  CountRecombEvents(simulated)
  CountRecombEvents(estimated)
  
```
And our estimated recombination rate could be calculated like this:
```{r}
  CountRecombEvents(estimated)/(10000*2)
  
  
```

Now testing different initial recombination rates
```{r}
  rates<-c(0.5,0.4,0.3,0.3,0.2,0.1,0.01,0.001)
  estimated.rate<-c()
  recomb.count<-c()
  index<-1;
  for(rate in rates){
    trans.p<-GetTransitionProb(rate); 
    v.2<-Viterbi3 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
    vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
    estimated<-rownames(trans.p)[vitrowmax.2]
    recomb.count[index]<-CountRecombEvents(estimated)
    estimated.rate[index]<- recomb.count[index]/(10000*2);
    index<-index+1
  }
  
  rates
  estimated.rate
  recomb.count
  

    
  plot(rates,estimated.rate,xlab="arbitrary recomb. rate given to transition matrix",ylab="estimated recomb. rate",main="r.rate estimations with different starting r.rate\n on one simulation with 10,000 loci and r=0.01")
  
```

#### Now we simulate wih 10,000 loci and r=0.05

```{r}
  recombination.rate<-0.05 #simulating with r=0.001
  trans.p<-GetTransitionProb(recombination.rate)
  child1.chromatid.mom<-SimulateChromatid(10000,recombination.rate) 
  child1.chromatid.dad<-SimulateChromatid(10000,recombination.rate)
  simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
  obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
  
```

Now testing different initial recombination rates
```{r}
  rates<-c(0.5,0.4,0.3,0.3,0.2,0.1,0.01,0.001)
  estimated.rate<-c()
  recomb.count<-c()
  index<-1;
  for(rate in rates){
    trans.p<-GetTransitionProb(rate);
    v.2<-Viterbi3 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
    vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
    estimated<-rownames(trans.p)[vitrowmax.2]
    recomb.count[index]<-CountRecombEvents(estimated)
    estimated.rate[index]<- recomb.count[index]/(10000*2);
    index<-index+1
  }
  
  rates
  estimated.rate
  recomb.count
  

    
  plot(rates,estimated.rate,xlab="arbitrary recomb. rate given to transition matrix",ylab="estimated recomb. rate",main="r.rate estimations with different starting r.rate\n on one simulation with 10,000 loci and r=0.05")
  
```

#### Now we simulate wih 10,000 loci and r=0.1

```{r}
  recombination.rate<-0.1 #simulating with r=0.001
  trans.p<-GetTransitionProb(recombination.rate)
  child1.chromatid.mom<-SimulateChromatid(10000,recombination.rate) 
  child1.chromatid.dad<-SimulateChromatid(10000,recombination.rate)
  simulated<-paste("mom.",child1.chromatid.mom$parent.path,"/","dad.",child1.chromatid.dad$parent.path,sep="")
  obs<-child1.chromatid.mom$child+child1.chromatid.dad$child
  
```

Now testing different initial recombination rates
```{r}
  rates<-c(0.5,0.4,0.3,0.3,0.2,0.1,0.01,0.001)
  estimated.rate<-c()
  recomb.count<-c()
  index<-1;
  for(rate in rates){
    trans.p<-GetTransitionProb(rate);
    v.2<-Viterbi3 (obs,start.p, trans.p,child1.chromatid.mom,child1.chromatid.dad)
    vitrowmax.2 <- apply(v.2, 1, function(x) which.max(x))
    estimated<-rownames(trans.p)[vitrowmax.2]
    recomb.count[index]<-CountRecombEvents(estimated)
    estimated.rate[index]<- recomb.count[index]/(10000*2);
    index<-index+1
  }
  
  rates
  estimated.rate
  recomb.count
  

    
  plot(rates,estimated.rate,xlab="arbitrary recomb. rate given to transition matrix",ylab="estimated recomb. rate",main="r.rate estimations with different starting r.rate\n on one simulation with 10,000 loci and r=0.1")
  
```

#### Author
Luis Avila (lmavila@gmail.com)