9. Considering missing data
-----------------------------
#### Adjusting our model


```{r eval=TRUE, echo=FALSE}

source("SharedFunctions.R")


```



The emission probability matrix has the probabilities of observing O~i~ given 
a G~i~

We have updated it to handle missing data.


```{r}
GetEmissionProbWithMissingData<-function(mom.maternal,mom.paternal,dad.maternal,dad.paternal,prob.missing){
  row.1<-c(as.numeric(mom.maternal+dad.maternal==0),
           as.numeric(mom.maternal+dad.maternal==1),
           as.numeric(mom.maternal+dad.maternal==2))
  row.2<-c(as.numeric(mom.maternal+dad.paternal==0),
           as.numeric(mom.maternal+dad.paternal==1),
           as.numeric(mom.maternal+dad.paternal==2)) 
  row.3<-c(as.numeric(mom.paternal+dad.maternal==0),
           as.numeric(mom.paternal+dad.maternal==1),
           as.numeric(mom.paternal+dad.maternal==2))
  row.4<-c(as.numeric(mom.paternal+dad.paternal==0),
           as.numeric(mom.paternal+dad.paternal==1),
           as.numeric(mom.paternal+dad.paternal==2))      
  
  emission.matrix<-rbind(row.1,row.2,row.3,row.4)
  emission.matrix<-emission.matrix*(1-prob.missing)
  emission.matrix<-cbind(emission.matrix,rep(prob.missing,4))
  colnames(emission.matrix)<-c('0','1','2','3')
  rownames(emission.matrix)<-c('mom.maternal/dad.maternal','mom.maternal/dad.paternal',
                               'mom.paternal/dad.maternal','mom.paternal/dad.paternal')
  
  return(emission.matrix)
}


```

Now we can test it with the alleles from genotype_0, genotype_2, and genotype_15
```{r}
GetEmissionProbWithMissingData(0,0,0,0,0.3) #genotype_0
GetEmissionProbWithMissingData(0,0,1,0,0.3) #genotype_2
GetEmissionProbWithMissingData(1,1,1,1,0.3) #genotype_15
```
And we have updated our viterbi function to handle missing data in the observed genotype data
and also in the parental phasing.

```{r}
ViterbiWithMissingData <- function(obs,start.p, trans.p,mom.chromatid,dad.chromatid,prob.missing) {
  #initializing
  v <- matrix(NA, nr=length(obs), nc=dim(trans.p)[1])
  if(max(mom.chromatid$maternal[1],
         mom.chromatid$paternal[1],
         dad.chromatid$maternal[1],
         dad.chromatid$paternal[1])>1){
    # if there is any error on the parental phasing
    # force an emission matrix ab initio
    emit.p<-GetEmissionProbWithMissingData(0,0,0,0,prob.missing)
  } else {
    emit.p<-GetEmissionProbWithMissingData(mom.chromatid$maternal[1],
                            mom.chromatid$paternal[1],
                            dad.chromatid$maternal[1],
                            dad.chromatid$paternal[1],prob.missing)
  }
  
  emit.p<-emit.p+10^(-12) 
  #print(emit.p)
  old.emit.pe<-emit.p
  for(i in 1:4){
    v[1,i]=log(start.p[i],2)+log(emit.p[i,obs[1]+1],2)
  }
  
  for(i in 2:length(obs)) {# from observation 2 to t
   if(max(mom.chromatid$maternal[i],
          mom.chromatid$paternal[i],
          dad.chromatid$maternal[i],
          dad.chromatid$paternal[i])>1){
      # if there is any error on the parental phasing
      #  keep old emission matrix
    emit.p<-old.emit.pe
    } else {
      emit.p<-GetEmissionProbWithMissingData(mom.chromatid$maternal[i],
                              mom.chromatid$paternal[i],
                              dad.chromatid$maternal[i],
                              dad.chromatid$paternal[i],prob.missing)
      old.emit.pe<-emit.p
    }
   # print(emit.p++10^(-12))
    for (l in 1:dim(trans.p)[1]) { #from state 1 to N (8 states)
      v[i,l] <- log(emit.p[l,obs[i]+1]+10^(-12),2) + max(v[(i-1),] + log( trans.p[l,],2))
    }
  }   
  return(v) 
}
```

#### Author
Luis Avila (lmavila@gmail.com)